\mytitle

### 2015-08-02 MSc_diss.tex
Changes in objectives and aims

\section{Introduction} \label{sec:intro}
The Earth’s atmosphere is the recipient of a vast range of chemical compounds emitted by natural and anthropogenic sources. The oxidation of these compounds is the main process taking place in the atmosphere that keeps their concentrations at levels that do not distort the chemical balance of the atmosphere \citep{Prinn2003}. In other words, oxidation is a removal 'cleansing' mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the oxidizing capacity of the atmosphere. This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.

\subsection{Tropospheric ozone}
Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere. Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the 'cleansing' capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}. On the other hand, being a source of main pollutant removal agents, ozone is an air pollutant itself. Long exposure to its high concentrations causes respiratory problems in humans and damage to sensitive plant species \citep{Fowler2008}. Last but not least characteristic of tropospheric ozone is that it is an important greenhouse gas with a current estimated radiative forcing of  $0.40\pm 0.20~Wm^{–2}$ which is approximately one fifth of the $CO_2$ radiative forcing \citep{Hartmann2013,Myhre2013}. Due to these reasons a considerable interest in quantifying tropospheric ozone concentrations, budget and trends exists for more than a 100 years \citep{Becker2004}.

@@

@@

According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system. For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}. Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as: 
\begin{equation}
\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
\end{equation}
where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$. In other words, chemical loss of 1 carbon bond produces 1 peroxy radical:
\begin{itemize}
\centering
\item[] $OH + {CH_4}|_{4 bonds} \rightarrow {CH_3O_2}|_{3 bonds} + H_2O$
\item[] 4 carbon bonds = 3 carbon bonds, 1 peroxy radical
\end{itemize}
Loss of 1 carbon bond via photolysis produces 2 peroxy radicals:
\begin{itemize}
\centering
\item[] ${HCHO}|_{2bonds} + hv \rightarrow H + {HCO}|_{1bond}$
\item[] ${HCO}|_{1bond} + O_2 \rightarrow HO_2 + {CO}|_{1bond}$
\item[] $H + O_2 \xrightarrow{M} HO_2$
\item[] $horizontal line$
\item[] 2 carbon bonds = 1 carbon bond, 2 peroxy radicals
\end{itemize}
or no peroxy radicals because of dry or wet deposition:
\begin{itemize}
\centering
\item[] $HCHO \xrightarrow{deposition}$
\item[] 2 carbon bonds = 2 carbon bonds, 0 peroxy radicals
\end{itemize}

@@

For link at the bottom? The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}. It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark.
\subsection{Motivation}
\citep{Sommariva2008}
[Typically, chemical mechanisms do not contain
enough information about the various formation routes of secondary organic compounds, leading to the underestimation of their concentrations, with possible consequences on the estimates of ozone production in remote regions, or to ineffective ozone-control policies. In this work, a quasi-explicit chemical model based upon the Master Chemical Mechanism (MCM) has been used to study in detail the photochemistry of a number of alkyl nitrates and determine quantitatively their formation processes. This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]

\section{Overall objective and specific aims}
\subsection{Overall objective}
The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
\subsection{Specific aims}
Specific aims are:
\begin{itemize}
\item to investigate of the relationship between ozone production and carbon bond emissions (see Data and Methods);
\item to analyse the impact of inclusion of alkyl nitrates chemistry on this relationship;
\item to analyse everything mentioned above in various NOx regimes.
\end{itemize}

\section{Literature review} \label{sec:data}
\subsubsection{Observational studies}

>>>>

\section{Introduction} \label{sec:intro}
The Earth’s atmosphere is the recipient of a vast range of chemical compounds emitted by natural and anthropogenic sources. The oxidation of these compounds is the main process taking place in the atmosphere that keeps their concentrations at levels that do not distort the chemical balance of the atmosphere \citep{Prinn2003}. In other words, oxidation is a removal 'cleansing' mechanism for these compounds (various pollutants and greenhouse gases), and its rate is often referred to as the oxidizing capacity of the atmosphere. This capacity depends mainly on two factors: climate and the global burden of the principal oxidants such as ozone ($O_3$), the hydroxyl radical ($OH$) and hydrogen peroxide ($H_2O_2$) \citep{Prinn2003,Thompson1992}.
\subsection{Tropospheric ozone}
Tropospheric ozone ($O_3$) plays an important role in the chemical and thermal balance of the atmosphere. Although it accounts only for about 10\% of the total ozone, it is a primary precursor to the formation of hydroxyl radicals ($OH$) which in association with hydrogen peroxide ($H_2O_2$) determines the 'cleansing' capacity of the atmosphere \citep{Prinn2003,Tarasick2008,Thompson1992}. On the other hand, being a source of main pollutant removal agents, ozone is an air pollutant itself. Long exposure to its high concentrations causes respiratory problems in humans and damage to sensitive plant species \citep{Fowler2008}. Last but not least characteristic of tropospheric ozone is that it is an important greenhouse gas with a current estimated radiative forcing of  $0.40\pm 0.20~Wm^{–2}$ which is approximately one fifth of the $CO_2$ radiative forcing \citep{Hartmann2013,Myhre2013}. Due to these reasons a considerable interest in quantifying tropospheric ozone concentrations, budget and trends exists for more than a 100 years \citep{Becker2004}.

@@

MAKE GRAPHIC FROM TABLE ABOUT O3 BUDGETS!

@@

According to \citep{Evans2014} (personal communication), the $O_3$ production efficiency can be estimated via calculation of the number of carbon bonds emitted into the system. For example, if the system contains one molecule of methane, $CH_4$, as 'a carbon bond reservoir', it means that this system has 4 carbon bonds (4 C-H), if  instead of methane it contains one molecule of ethane, $C_2H_6$, the system has 7 carbon bonds (6 C-H bonds and 1 C-C), and so on \citep{Edwards2013}. Assuming that every carbon bond is broken chemically and every organic peroxy radical ($RO_2$) reacts with $NO$, the overall efficiency of carbon bonds conversion to $O_3$, $\epsilon$, can be calculated as: 
\begin{equation}\label{eq:epsilon}
\epsilon = \dfrac{P_{O_3}}{E_{C_{bons}}} = \chi\cdot\upsilon
\end{equation}
where $P_{O_3}$ – the rate of $O_3$ production (mol yr-1), $E_{C_{bons}}$  – the rate of carbon bonds emission (mol yr-1), $\chi$ – ratio between the rate of carbon bonds loss to form $RO_2$ and the loss rate of all carbon bonds, $\upsilon$ – ratio between the rate of $RO_2$ loss in reaction with $NO$ resulting in $NO$-to-$NO_2$ conversion and the loss rate of all $RO_2$. In other words, chemical loss of 1 carbon bond produces 1 peroxy radical:
\begin{itemize}
\centering
\item[] $OH + {CH_4}|_{4 bonds} \rightarrow {CH_3O_2}|_{3 bonds} + H_2O$
\item[] 4 carbon bonds = 3 carbon bonds, 1 peroxy radical
\end{itemize}
Loss of 1 carbon bond via photolysis produces 2 peroxy radicals:
\begin{itemize}
\centering
\item[] ${HCHO}|_{2bonds} + hv \rightarrow H + {HCO}|_{1bond}$
\item[] ${HCO}|_{1bond} + O_2 \rightarrow HO_2 + {CO}|_{1bond}$
\item[] $H + O_2 \xrightarrow{M} HO_2$
\item[] $horizontal line$
\item[] 2 carbon bonds = 1 carbon bond, 2 peroxy radicals
\end{itemize}
or no peroxy radicals because of dry or wet deposition:
\begin{itemize}
\centering
\item[] $HCHO \xrightarrow{deposition}$
\item[] 2 carbon bonds = 2 carbon bonds, 0 peroxy radicals
\end{itemize}

@@

For link at the bottom? The same kind of approach is used in the Common Representative Intermediates (CRI) mechanism for VOC degradation \citep{Jenkin2002, Jenkin2008}. It is a reduced chemical mechanism that describes $O_3$ formation from the tropospheric degradation of methane and 115 emitted non-methane hydrocarbons and oxygenated volatile organic compounds developed using the Master Chemical Mechanism version 3.1 (MCM v3.1) as a reference benchmark
\section{Overall objective and specific aims}
As the reader might anticipateit is mentioned in section \ref{sec:lit_review}
\citep{Sommariva2008}
[Typically, chemical mechanisms do not contain
enough information about the various formation routes of secondary organic compounds, leading to the underestimation of their concentrations, with possible consequences on the estimates of ozone production in remote regions, or to ineffective ozone-control policies. In this work, a quasi-explicit chemical model based upon the Master Chemical Mechanism (MCM) has been used to study in detail the photochemistry of a number of alkyl nitrates and determine quantitatively their formation processes. This kind of study shows the potential of using mechanisms with a high level of chemical detail to study the formation of pollutants in the atmosphere (Madronich, 2006).]
\subsection{Overall objective}
The overall purpose of this research is to investigate the impact of alkyl nitrate chemistry on ozone production efficiency.
\subsection{Specific aims}
Specific aims are to investigate:
\begin{itemize}
\item[(1)] the relationship between ozone production and carbon bond emissions;
\item[(2)] the impact of inclusion of alkyl nitrates chemistry on this relationship;
\item[(3)] how (1) and (2) vary in different $NO_x$ and $VOC$ regimes;
\item[(4)] build different version of the box model;
\item[(5)] to evaluate the results.
\end{itemize}

\section{Literature review} \label{sec:lit_review}
\subsubsection{Observational studies}
